---
title: "Binomial of sum of Gaussians versus sum of binomial of Gaussians"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r}
library(ggplot2)
library(magrittr)
```

# IID latent field

**Without aggregation**: $y = \sum_{i = 1}^n y_i$, $y_i \sim \text{Bin}(m, \rho_i)$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)$

**With aggregation**: $y \sim \text{Bin}(nm, \bar \rho)$, $\bar \rho = \frac{1}{n} \sum_{i = 1}^n \rho_i$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)$

```{r}
experiment <- function(nsim, n, phi_mean, phi_sd, y_sample_size) {
  phi <- rnorm(n, phi_mean, phi_sd)
  rho <- plogis(phi)
  m <- rep(y_sample_size, n)
  
  # Without aggregation
  y <- replicate(nsim, rbinom(n, m, rho), simplify = FALSE)
  y <- lapply(y, sum)
  
  # With aggregation (at the level of the latent field)
  y_agg <- replicate(nsim, rbinom(1, sum(m), weighted.mean(rho)), simplify = FALSE)
  
  return(list(y = unlist(y), y_agg = unlist(y_agg), phi = phi, rho = rho))
}
```

## $\rho$ small setting (disease)

```{r}
nsim <- 500
n <- 36
result <- experiment(nsim, n, -2.5, 0.5, 30)
```

This is what the underlying latent field looks like for these settings:

```{r}
plot(result$rho)
```

And then the difference between the truth and the aggregated approximation:

```{r}
hist_fun <- function(result) {
  data.frame(type = c(rep("No aggregation (truth)", nsim), rep("Aggregation (approximation)", nsim)),
             y    = c(result$y, result$y_agg)) %>%
  ggplot() +
  geom_histogram(aes(x = y, fill = type), alpha = 0.5, position = "identity")
}

hist_fun(result)
```

```{r}
sd(result$y)
sd(result$y_agg)
sd(result$y_agg) / sd(result$y) # Ratio
```

## $\rho \approx 0.5$ setting (elections)

As before but with different `phi_mean` settings:

```{r}
result_half <- experiment(nsim, n, 0, 0.5, 30)
```

```{r}
plot(result_half$rho)
```

```{r}
hist_fun(result_half)
```

```{r}
sd(result_half$y)
sd(result_half$y_agg)
sd(result_half$y_agg) / sd(result_half$y) # Ratio
```


# Conclusions, things to add

* Standard deviation of aggregation overestimates true standard deviation, by a larger ratio in the smaller $\rho$ setting than the larger $\rho$ setting
* Should test latent field with spatial structure e.g. ICAR latent field and so on
* What is the effect of varying `sd_phi`? What is the effect of varying `y_sample_size`?